\section{Method}
\label{sec:method}

\subsection{Preprocessing}

All articles pass through a shared preprocessing step before any modeling. Each article is split on whitespace, then each token is:

\begin{enumerate}
    \item lowercased,
    \item filtered to retain only tokens matching the pattern \texttt{[\^{}a-zA-Z0-9]} (i.e., alphabetic-initial tokens),
    \item filtered to remove tokens shorter than 2 characters,
    \item optionally filtered against an English stopword list (NLTK \texttt{stopwords.words("english")}).
\end{enumerate}

Stopword removal is applied for genre classification and keyword scoring, where function words carry no topical signal. It is \emph{not} applied for sentiment classification, where negation words (``not,'' ``never'') carry meaningful polarity information.

Unlike many production NLP pipelines, we deliberately avoid stemming or lemmatization. Retaining full word forms keeps the vocabulary human-readable in keyword output and avoids the risk of conflating morphologically similar but semantically distinct terms (e.g., \textit{bank} and \textit{banking} in a business context).

\subsection{Feature Extraction}

\paragraph{Bag-of-Words for classification.}
Both the genre and sentiment classifiers represent documents as bags of words — unordered multisets of tokens after preprocessing. When \texttt{binarize=True} (disabled by default), each token type is counted at most once per document, converting the representation to a binary presence/absence feature vector. In practice we found standard token counts to perform comparably to binarized counts on this corpus.

\paragraph{TF-IDF for keyword ranking.}
For keyword extraction we build an inverted index over all 2,225 articles and compute TF-IDF weights using the standard log-frequency formulation:

\[
\text{tf-idf}(t, d) = (1 + \log_{10} \text{tf}_{t,d}) \times \log_{10}\frac{N}{\text{df}_t}
\]

where $\text{tf}_{t,d}$ is the raw term frequency in document $d$, $N$ is the total number of documents, and $\text{df}_t$ is the number of documents containing term $t$. For articles not seen during index construction (e.g., live RSS articles), we reuse the corpus IDF values and compute TF-IDF on the fly.

\paragraph{POS-tag filtering.}
To produce more readable keyword lists, we additionally filter terms by their NLTK part-of-speech tag, retaining only nouns (NN, NNS, NNP, NNPS) and adjectives (JJ, JJR, JJS). This removes verbs, determiners, and other grammatical function words that might receive high TF-IDF scores in short snippets.

% -----------------------------------------------------------------------
% TODO: Consider a two-column figure comparing raw TF-IDF keywords
%   vs. POS-filtered keywords for a sample article to illustrate
%   the benefit of the POS filter.
% -----------------------------------------------------------------------

\subsection{Named Entity Recognition}

Named entities are extracted using NLTK's \texttt{ne\_chunk} function \cite{bird2009natural}, which applies a binary Maximum Entropy NE chunker on top of POS-tagged tokens. We extract entities of types PERSON, ORGANIZATION, GPE (geopolitical entity), GSP (geographical/social/political), and FACILITY. Each article's entities are de-duplicated and grouped by type before display.

\subsection{Genre Classifier}

We implement Multinomial Naive Bayes from scratch following the derivation in \citet{mccallum1998comparison}. For a document $d$ with token list $\mathbf{w}$ and class set $\mathcal{C}$, the predicted class is:

\[
\hat{c} = \arg\max_{c \in \mathcal{C}} \left[ \log P(c) + \sum_{w \in \mathbf{w}} \log P(w \mid c) \right]
\]

Parameters $P(c)$ are estimated by relative class frequency in the training set. Conditional probabilities $P(w \mid c)$ are estimated with Laplace (add-$\alpha$) smoothing:

\[
P(w \mid c) = \frac{\text{count}(w, c) + \alpha}{\sum_{w'} \text{count}(w', c) + \alpha \cdot |V|}
\]

where $|V|$ is the vocabulary size and $\alpha = 1.0$ by default. All computations are done in log-space to avoid floating-point underflow. At inference time, words outside the training vocabulary are silently skipped (their log-probability contribution is zero).

\subsection{Sentiment Analysis}

Sentiment classification uses two independent systems whose outputs are compared:

\begin{itemize}
    \item \textbf{VADER} \cite{hutto2014vader}: a rule-based analyzer using a curated sentiment lexicon with heuristics for capitalization, punctuation, and valence shifters (negation, degree adverbs). It produces a compound score in $[-1, +1]$; we threshold at $\pm 0.05$ to assign positive/negative/neutral labels.

    \item \textbf{NB Sentiment}: the same Multinomial Naive Bayes architecture used for genre classification, trained on BBC articles pseudo-labeled by VADER (neutral articles excluded). Training on in-domain text means the model has seen the vocabulary and syntactic patterns of news writing, which we hypothesize should produce more reliable sentiment predictions than a model trained on movie reviews or social media.
\end{itemize}

% -----------------------------------------------------------------------
% TODO: Insert a figure illustrating VADER's compound score distribution
%   over the BBC corpus — a histogram showing the tri-modal structure
%   (negative / neutral / positive buckets) would clearly motivate
%   dropping neutral articles from sentiment training.
% -----------------------------------------------------------------------

\subsection{Why NLTK?}

NLTK \cite{bird2009natural} was chosen as the primary NLP toolkit for several reasons. First, it provides all the components needed in a single, well-documented library: tokenization, POS tagging, named-entity chunking, the VADER sentiment analyzer, and access to standard corpora. Second, its APIs closely match the conceptual descriptions in the \textit{NLTK Book} \cite{bird2009natural}, making the implementation transparent and educational. Third, its rule-based taggers and chunkers are deterministic and require no additional trained model downloads beyond the standard data packages, making the project straightforward to set up. Where statistical learning is required (the genre and sentiment classifiers), we use our own implementation rather than wrapping an NLTK class, to demonstrate that the core algorithm is understandable and not a black box.
